---
title: 'Raw dataset construction for the project Redundancy of reproductive traits'
author: "Erick Calderon-Morales"
date: '2022'
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
					  fig.width = 11, fig.height = 7)
```

```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      
                      # Save a pdf_copy?
                      #dev            = c( "png", "pdf"),
                      # Include code?
                      echo           = TRUE,                      
                      error          = FALSE,
                      fig.align      = "center",
                      # Path where figures are going to be store pdf single figures
                      fig.path       = paste0("./notebooks/figures_rawdataset", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```

```{r libaries, message=FALSE, warning=FALSE, cache=FALSE}
library(janitor)
library(dplyr)
# For tables
library(reactable)
# For pivot function
library(tidyr)
# For reading html files
library(rvest)
# For pluck function
library(purrr)
# For read xls files
library(readxl)
# For working with names
library(stringr)
# For converting pdf to data
library(tabulizer) 
# For adding columns
library(tibble)
# For wood density estimation
library(BIOMASS)
```


# Species list
## Full species list 

```{r}
getwd()
# Load species list that includes morpho species
source("./scripts/script_full_species_list.R")

# Spcode with the first 4 characters from the genus and the first 2 from species
data_species_full_list <-
    data_species_full_list %>%
        # This step is done for when mophospecies is sp get it and if it is sp## 
        # get that 4 characters
        mutate(gen4 = str_extract(genero, "^.{4}"),
                sp3 = if_else(str_length(especie) == 2, str_extract(especie, "^.{2}"),
                             if_else((str_length(especie) > 2) & (str_length(especie) <= 4) & (str_detect(especie, "^sp")),
                                     str_extract(especie, "^.{4}"), str_extract(especie, "^.{3}")))
               ) %>%
        unite(spcode_4_3, c("gen4","sp3"), sep = "") %>% 
        arrange(spcode)

reactable(data_species_full_list)
```

```{r}
data_names_spcode <- 
    data_species_full_list %>% 
        mutate(genero = str_to_title(genero)) %>% 
    
        # Create name_submitted column
        unite(name_submitted, c(genero, especie),sep = " ", remove = F) %>% 
        
        # Remove morphospecies
        filter(!(str_detect(especie, "^sp") & (str_length(especie) <= 4))) %>% 
        
        select(-c(genero,especie,familia))
```

```{r}
# Load TNRS species list 
tnrs_species_list <- read.csv("./data/raw_data/tnrs_names_short.csv", header = T) %>% 
    clean_names()
```

```{r}
# Include spcodes in TNRS file
tnrs_species_list <- 
    tnrs_species_list %>% 
        select(-id) %>% 
        inner_join(., data_names_spcode, by = "name_submitted") %>% 
        select(spcode,spcode_4_3, everything())
```

## Full species list without morpho-species

__The accepted_species columns were obtained from the TRNS__ 

```{r}
species_list <- 
    tnrs_species_list %>% 
        # Species' name manually changed
        mutate(accepted_species = case_when(
            name_submitted == "Billia colombiana" ~ "Putzeysia rosea",
            name_submitted == "Hyeronima oblonga" ~ "Stilaginella oblonga",
            TRUE ~ accepted_species)) %>%  

        select(spcode, spcode_4_3, name_submitted, taxonomic_status, 
               accepted_species) %>% 
        arrange(name_submitted) %>% 
    
        # In this list Brosimum panamense is treat as a different species and is 
        # not. Removed
        filter(!name_submitted == "Brosimum panamense",
               !name_submitted == "Hirtella media") %>%  
        
        select(-taxonomic_status)

# Total should be 190 rows
reactable(species_list)
```

# Reproductive traits datasets

##  Traits from Salgado

### Change the Abbreviations 

```{r echo = FALSE}
salgado_reproductive_traits_abbr <- 
    tribble(
     ~trait, ~type, ~original_abbr,~new_abbr,
      "Dispersion syndrome","Wind (anemocoria)", "W","dwin", 
      "Dispersion syndrome","Water (hidrocoria)", "H","dhid",
      "Dispersion syndrome","Animals (Flying animals)", "V","dani",
      "Dispersion syndrome","Animals (Non-flying animals)","NV" ,"dani",
      "Dispersion syndrome","Autorcoria", "A","daut",
      "Dispersion syndrome","Species with several syndromes", NA,"dvar",
      
      "Sexual System","Dioecious", "D", "ssdio",
      "Sexual System","Monoecious", "M", "ssmon",
      "Sexual System","Hermaphroditic", "H", "ssher",
      "Sexual System","Poligama", "P" , "Not taken into account",
      
      "Pollination syndrome","Wind", "W" , "pwin",
      "Pollination syndrome","Insects", "I", "pins",
      "Pollination syndrome","Birds", "A", "pbir",
      "Pollination syndrome","Zoocoria (No voladora)", "MA", "pmam",
      "Pollination syndrome","Species with several dispersal syndromes", NA, "pvar",
     )
  
reactable(
  salgado_reproductive_traits_abbr,
  groupBy = "trait",
  columns = list(trait = colDef(minWidth = 165),
                 type = colDef(minWidth  = 125),
                 original_abbr = colDef(minWidth = 100)),
  rowStyle = JS("function(rowInfo) {
    if (rowInfo.level > 0) {
      return { background: '#eee', borderLeft: '2px solid #ffa62d' }
    } else {
      return { borderLeft: '2px solid transparent' }
    }
  }"),
)
```

```{r}
# Load Original file
raw_salgado_traits <- read_xls("./data/raw_data/traits/response_traits/original_raw_data_files/Costa Rica rainforest trees 2 Salgado_original.xls") %>% 
    clean_names()
```

### Species with traits in Salgado dataset

```{r}
data_salgado_traits <- 
    raw_salgado_traits %>%
        select(-c(familia, autoridad)) %>% 
        unite("name_submitted", genero, especie, sep = " ", remove = F) %>%  
        
        # Species with traits in Salgado dataset 133
        inner_join(species_list, . , by = "name_submitted") %>% 
        select(accepted_species, everything(),
               -c(estrato,tasacrecimiento,rep_vegetativa))  %>% 
        
        # Rename columns
        rename(dispersal_syndrome = "diseminacion",
               sexual_system = "sist_sexual",
               pollination_syndrome = "polinizacion") %>% 
        
        # Remove reference numbers, parentheses and comas
        mutate(across(where(is.character), str_replace_all, pattern = "[:digit:]|[(,)]",
                      replacement = ""))  %>% 
        
        # Rename levels in dispersion column
        mutate(dispersal_syndrome_modified = case_when(
            # Wind
            dispersal_syndrome == "W"  ~ "dwin",
            # Hidrocoria
            dispersal_syndrome == "H"  ~ "dhid",
            # Convert Flying animals Non-flying animals to just animals
            dispersal_syndrome == "V" | dispersal_syndrome == "NV"  ~ "dani",
            # Autocoria
            dispersal_syndrome == "A" ~ "daut",
            str_detect(dispersal_syndrome, "-")  ~ "dvar",
            TRUE ~ dispersal_syndrome)) %>%
        
        # Rename levels in sexual_system column
        mutate(sexual_system_modified = case_when(
            # Dioecious
            sexual_system == "D"  ~ "ssdio",
            # Monoecious
            sexual_system == "M"  ~ "ssmon",
            # Hermaphroditic
            sexual_system == "H" ~ "ssher",
            TRUE ~ sexual_system)) %>% 
        
        # Rename levels in pollination column
        mutate(pollination_syndrome_modified = case_when(
            # Wind
            pollination_syndrome == "W"  ~ "pwin",
            # Insects
            pollination_syndrome == "I"  ~ "pins",
            # Birds
            pollination_syndrome == "A" ~ "pbir",
            # Mammals
            pollination_syndrome == "MA" ~ "pmam",
            # Various 
            str_detect(pollination_syndrome, "-")  ~ "pvar",
            TRUE ~ pollination_syndrome))  %>% 
        
        # Add author
        add_column(source = "bsalgado dataset") %>% 
    
        # Sort columns
        select(sort(current_vars())) %>% 
        select(spcode, spcode_4_3, accepted_species, name_submitted, 
               genero, especie, everything()) %>% 
        arrange(genero, especie)

```

```{r}
# Species traits
# This step is done for listing the species with traits 

abbr_change_salagado <- 
    data_salgado_traits %>% 
        select(-c(spcode, spcode_4_3, genero,especie, source, name_submitted)) %>%   
        unite(dispersal_syndrome, c(2:3), sep = "/")  %>%  
        unite(pollination_syndrome, c(3:4), sep = "/") %>% 
        unite(sexual_system, c(4:5), sep = "/")  %>% 
        
        # Pivot for compare old abbr with new abbr
        pivot_longer(.,
                     cols = 2:4,
                     names_to = "trait",
                     values_to = "abbr") %>% 
        separate(abbr,c("abbr_original","abbr_new"), sep = "/")
    
reactable(abbr_change_salagado,
          groupBy = "accepted_species",
          columns = list(trait = colDef(minWidth = 165),
                         abbr_original = colDef(minWidth  = 125),
                         abbr_new = colDef(minWidth = 100)),
          rowStyle = JS("function(rowInfo) {
    if (rowInfo.level > 0) {
      return { background: '#eee', borderLeft: '2px solid #ffa62d' }
    } else {
      return { borderLeft: '2px solid transparent' }
    }
  }"),
)
```

```{r}
# Species with no traits in Salgado dataset
species_with_no_traits_58 <- 
    raw_salgado_traits %>% 
        select(-c(familia, autoridad)) %>% 
        unite("name_submitted", genero, especie, sep = " ") %>% 
    
        # Species with NO traits in Salgado dataset
        anti_join(species_list, . , by = "name_submitted") %>% 
        select(spcode, spcode_4_3, accepted_species, everything())
```

## Traits from Chazdon et al

### Change in Abbreviations

### Species with traits in Chazdon et al


```{r}
chazdon_reproductive_traits_abbr <- 
    tribble(
     ~trait, ~type, ~original_abbr,~new_abbr,
      "Dispersion syndrome","Animals", "A","dani", 
      "Dispersion syndrome","Explosive", "E","daut",
      "Dispersion syndrome","Wind", "W","dwin",
      "Dispersion syndrome", "Gravity", "G", "daut", 
      "Dispersion syndrome", "Unknown", "UN", "NA",
      
      "Sexual System","Dioecious","D","ssdio",
      "Sexual System","Monoecious", "M","ssmon",
      "Sexual System","Hermaphroditic", "H","ssher",
      "Sexual System", "Unknown", "UN", "NA",
     
      "Pollination syndrome", "Wind","W" ,"pwin",
      "Pollination syndrome", "Insects", "I","pins",
      "Pollination syndrome", "Birds", "Hb","pbir",
      "Pollination syndrome", "Mammal", "MA","pmam",
      "Pollination syndrome", "Unknown", "UN", "NA"
     )
  
reactable(
  chazdon_reproductive_traits_abbr,
  groupBy = "trait",
  width = "750",
  height = "750",
  columns = list(trait = colDef(minWidth = 165),
                 type = colDef(minWidth  = 125),
                 original_abbr = colDef(minWidth = 100)),
  rowStyle = JS("function(rowInfo) {
    if (rowInfo.level > 0) {
      return { background: '#eee', borderLeft: '2px solid #ffa62d' }
    } else {
      return { borderLeft: '2px solid transparent' }
    }
  }"),
) 
```

```{r}
# Load html original file
raw_chazdon_traits <- 
    read_html("./data/raw_data/traits/response_traits/original_raw_data_files/reproductive traits for 366 species in 10 wet tropical forests_chazdon2003.html") %>% 
        html_table(header = T) %>% 
        purrr::pluck(1)  %>% 
        clean_names() 
```

```{r}
data_chazdon_traits <- 
    raw_chazdon_traits %>% 
        
        # Remove column
        select(-c(life_form, family)) %>% 
        
        # Remove \n\t and dots from strings 
        mutate(across(2:4, str_replace_all, pattern = "[\n\t]|[:digit:]|[(,)]|[.]|[ ]",
                      replacement = ""))  %>% 
        
        # Step for removing authority
        separate(species, c("genero", "especie"))    %>% 
        
        # Rename levels in dispersion column
        mutate(dispersal_syndrome_modified = case_when(
            # Wind
            dispersal_syndrome == "W"  ~ "dwin",
            # Animals
            dispersal_syndrome == "A" ~ "dani",
            # Autocoria
            dispersal_syndrome == "E" | dispersal_syndrome == "G"  ~ "daut",
            # Unknown
            dispersal_syndrome == "UN" ~ "NA",
            TRUE ~ dispersal_syndrome)) %>% 
    
        # Rename levels in sexual_system column
        mutate(sexual_system_modified = case_when(
            # Dioecious
            sexual_system == "D"  ~ "ssdio",
            # Monoecious
            sexual_system == "M"  ~ "ssmon",
            # Hermaphroditic
            sexual_system == "H" ~ "ssher",
            # Unknown
            sexual_system == "UN" ~ "NA",
            TRUE ~ sexual_system)) %>%
    
        # Rename levels in pollination column
        mutate(pollination_syndrome_modified = case_when(
            # Wind
            pollination_syndrome == "W"  ~ "pwin",
            # Insects
            pollination_syndrome == "I"  ~ "pins",
            # Birds
            pollination_syndrome == "HB" ~ "pbir",
            # Mammals
            pollination_syndrome == "MA" ~ "pmam",
            # Unknown
            pollination_syndrome == "UN" ~ "NA",
            TRUE ~ pollination_syndrome)) %>% 
    
        arrange(genero, especie) %>% 
        
        # Join columns for get full name use TNRS accepted species# Remove morphospecies
        filter(!(str_detect(especie, "^sp") & (str_length(especie) <= 4))) %>% 
        unite("species_name", genero, especie, sep = " ", remove = F) %>% 
        
        # Rename species names that are wrong
        mutate(species_name = case_when(
            species_name == "Cespedesia spathulata"  ~ "Cespedesia macrophylla",
            species_name == "Cojoba catenatum"  ~ "Cojoba catenata",
            species_name == "Inga thiboudiana" ~ "Inga thibaudiana",
            TRUE ~ species_name))  %>% 
        
        # Species with traits in Chazdon et al
        inner_join(species_with_no_traits_58, . , by = c("name_submitted" = "species_name" )) %>%  
            
        add_column(source = "https://doi.org/10.6084/m9.figshare.c.3309012.v1")  %>% 
            
        # Sort columns
        select(sort(current_vars())) %>% 
        select(spcode, spcode_4_3, accepted_species, name_submitted,
               genero, especie, everything()) %>% 
        arrange(genero, especie)

# nrow should be 12
```

```{r}
# Abbr
abbr_change_chazdon <- 
    data_chazdon_traits %>% 
        select(-c(spcode, spcode_4_3, genero,especie, source, name_submitted)) %>% 
        unite(dispersal_syndrome, c(2:3), sep = "/")  %>%  
        unite(pollination_syndrome, c(3:4), sep = "/") %>% 
        unite(sexual_system, c(4:5), sep = "/") %>%
        
        # Pivot for compare old abbr with new abbr
        pivot_longer(.,
                     cols = 2:4,
                     names_to = "trait",
                     values_to = "abbr") %>% 
        separate(abbr,c("abbr_original","abbr_new"), sep = "/")
    
reactable(abbr_change_chazdon,
          groupBy = "accepted_species",
          width = "1000",
          height = "750",
          columns = list(trait = colDef(minWidth = 100),
                         abbr_original = colDef(minWidth  = 75),
                         abbr_new = colDef(minWidth = 75)),
          rowStyle = JS("function(rowInfo) {
    if (rowInfo.level > 0) {
      return { background: '#eee', borderLeft: '2px solid #ffa62d' }
    } else {
      return { borderLeft: '2px solid transparent' }
    }
  }"),
)
```

```{r}
# Species with no traits
species_with_no_traits_46 <-
     data_chazdon_traits %>% 
         anti_join(species_with_no_traits_58, ., by = c("accepted_species" = "accepted_species")) %>% 
         arrange(accepted_species)
```

## Species identified by Nelson Zamora

```{r}
raw_zamora_traits <- 
    read_xlsx("./data/raw_data/traits/response_traits/original_raw_data_files/species_identified_NZamora_original.xlsx") %>% 
    clean_names() %>% 
    select(-familia)
```

### Species with traits

```{r}
data_zamora_traits <- 
    raw_zamora_traits %>%
    
        # Join columns for get full name
        unite("name_submitted", genero, especie, sep = " ", remove = F) %>%
        
        select(-tipo_de_dispersion_original) %>% 
    
        # Rename columns
        rename(dispersal_syndrome = "tipo_de_dispersion",
               sexual_system = "sistema_sexual",
               pollination_syndrome = "agente_polinizador")  %>% 
    
        # Rename levels in dispersion column
        mutate(dispersal_syndrome_modified = case_when(
            # Wind
            dispersal_syndrome == "W"  ~ "dwin",
            # Hidrocoria
            dispersal_syndrome == "H"  ~ "dhid",
            # Animals
            dispersal_syndrome == "A" | dispersal_syndrome == "NW" | dispersal_syndrome == "AN"  ~ "dani", 
            # Autocoria
            dispersal_syndrome == "E" ~ "daut",
            str_detect(dispersal_syndrome, ",")  ~ "dvar",
            TRUE ~ dispersal_syndrome)) %>%
    
        # Rename levels in sexual_system column
        mutate(sexual_system_modified = case_when(
            # Dioecious
            sexual_system == "D"  ~ "ssdio",
            # Monoecious
            sexual_system == "M"  ~ "ssmon",
            # Hermaphroditic
            str_detect(sexual_system, pattern = "[bisexual]|[H]") ~ "ssher",
            TRUE ~ sexual_system)) %>% 
    
        # Rename levels in pollination column
        mutate(pollination_syndrome_modified = case_when(
            # Wind
            pollination_syndrome == "W"  ~ "pwin",
            # Mammals
            str_detect(pollination_syndrome, pattern = "murc") ~ "pmam",
            # Insects
            str_detect(pollination_syndrome, pattern = "[bee]|[insec]|[I]")  ~ "pins",
            # Birds
            pollination_syndrome == "HB" ~ "pbir",
            TRUE ~ pollination_syndrome))  %>%
    
        arrange(name_submitted) %>%  
        inner_join(species_with_no_traits_46 ,. , by = "name_submitted") %>% 
    
        add_column(source = "Nelson Zamora, personal communication") %>% 
        
        # Sort columns
        select(sort(current_vars())) %>%  
        select(spcode, spcode_4_3, accepted_species, name_submitted,
               genero, especie, everything(),
               # Remove columns
               -c(info)) %>% 
        arrange(genero, especie)

#Bisexual: each flower of each individual has both male and female structures
# hermaphroditism, the condition of having both male and female reproductive organs.
# bisexual == herma
```

```{r}
# Abbr
abbr_change_zamora <- 
    data_zamora_traits %>% 
        select(-c(spcode, spcode_4_3, genero,especie, source, name_submitted)) %>% 
        unite(dispersal_syndrome, c(2:3), sep = "/")  %>% 
        unite(pollination_syndrome, c(3:4), sep = "/")  %>% 
        unite(sexual_system, c(4:5), sep = "/") %>% 
        
        # Pivot for compare old abbr with new abbr
        pivot_longer(.,
                     cols = 2:4,
                     names_to = "trait",
                     values_to = "abbr") %>% 
        separate(abbr,c("abbr_original","abbr_new"), sep = "/")
    
reactable(abbr_change_zamora,
          groupBy = "accepted_species",
          width = "1000",
          height = "750",
          columns = list(trait = colDef(minWidth = 100),
                         abbr_original = colDef(minWidth  = 75),
                         abbr_new = colDef(minWidth = 75)),
          rowStyle = JS("function(rowInfo) {
    if (rowInfo.level > 0) {
      return { background: '#eee', borderLeft: '2px solid #ffa62d' }
    } else {
      return { borderLeft: '2px solid transparent' }
    }
  }"),
)
```

```{r}
# Species with no traits
species_with_no_traits_13 <- 
    data_zamora_traits %>% 
    
    # Join columns for get full name
    unite("name_submitted", genero, especie, sep = " ") %>% 
    
    # Species with traits identyfied by Nelson Zamora
    anti_join(species_with_no_traits_46, ., by = "name_submitted")
```

## Species identified by Orlando Vargas

```{r cache = TRUE, pdf_conversion_to_data}
# raw_data_vargas <- 
#     extract_tables("./data/raw_data/response_traits/original_raw_data_files/lista_arboles_sindromes_OVR05.pdf", 
#                                    
#                                    # Read this pages
#                                    pages = c(1:14), 
#                                    method = "lattice")


# This step is done because I dont trust the tabulizer package 
#save(raw_data_vargas, file = "./data/raw_data/response_traits/raw_data_vargas.RData")
# Restore the object
load("./data/raw_data/traits/response_traits/raw_data_vargas.RData")


```

### Species with traits
```{r}
data_vargas_traits <- 
    map_dfr(raw_data_vargas, as_tibble) %>% 
        row_to_names(row_number = 1) %>% 
        clean_names()   %>% 
        
        # Column poli-viento not included
        # only three species Weinmannia pinnata,Eschweilera costaricensis 
        # Sorocea pubivena poli by wind
        rename( dis_mamifero = "x",
                dis_aves     = "x_2",
                dis_viento   = "x_3",
                dis_agua     = "x_4",
                gravedad_explosion = "x_5",
                comentarios_1 = "x_6",
                poli_aves = "x_7",
                poli_mamiferos = "x_8",
                poli_insectos = "x_9",
                poli_viento = "x_10",
                comentarios_2 = "x_11",
                sexual_system = "sistema_sexual_bisexual_monoica_dioica_unisexual_poligama_poligama_dioica") %>% 
        select(-c(autores, nombre_comun, comentarios_1,comentarios_2))  %>% 
        
        unite("name", genero, especie, sep = " ", remove = F) %>% 
        # Rename species names that are wrong
        mutate(name = case_when(
            name == "Tabebuia chrysantha??"  ~ "Tabebuia chrysantha",
            TRUE ~ name)) %>% 
        
        # Select the species that I need
        inner_join(species_with_no_traits_13 ,., 
                   by = c("name_submitted" = "name" ))  %>% 
    
        # Rename levels in dispersal columns  
        mutate(
            # Animals
            dis_mamifero = case_when(
                dis_mamifero == 1 ~ "dani",
                TRUE ~ dis_mamifero),
        
            dis_aves = case_when(
                dis_aves == 1 ~ "dani",
                TRUE ~ dis_aves),
            
            # Wind
            dis_viento = case_when(
                dis_viento == 1 ~ "dwin",
                TRUE ~ dis_viento),
            
            # Hidrocoria
            dis_agua = case_when(
                dis_agua == 1 ~ "dhid",
                TRUE ~ dis_agua),
            
            # Autocoria
            gravedad_explosion = case_when(
                gravedad_explosion == 1 ~ "daut",
                TRUE ~ gravedad_explosion)) %>% 
            
            # Rename levels in dispersal columns
            mutate(
              # Birds
              poli_aves = case_when(
                  poli_aves == 1 ~ "pbir",
                  TRUE ~ poli_aves),
              
              # Mammals
              poli_mamiferos = case_when(
                  poli_mamiferos == 1 ~ "pmam",
                  TRUE ~ poli_mamiferos),
              
              # Insects
              poli_insectos = case_when(
                  poli_insectos == 1 ~ "pins",
                  TRUE ~ poli_insectos),
              
              # Wind
              poli_viento = case_when(
                  poli_viento == 1 ~ "pwin",
                  TRUE ~ poli_viento))  %>% 
        
        # Create new colums
        unite(dispersal_syndrome_modified, 8:12, sep = "_")  %>%  
        unite(pollination_syndrome_modified, 9:12, sep = "_")  %>%  
    
        # Delete special characters
        mutate(across(8:ncol(.), str_replace_all, 
                      pattern = "__|[*]|[?]",replacement = "")) %>% 
        
        # Add category dvar for species with more than 1     
        mutate(dispersal_syndrome_modified = case_when(
            str_detect(dispersal_syndrome_modified, "_")  ~ "dvar",
            TRUE ~ dispersal_syndrome_modified)) %>%  
        
        # Remove underscore from pins
         mutate(across(9, str_replace_all, pattern = "_", replacement = "")) %>% 
        
        # Remove columns
        select(-c(familia))  %>% 
    
        # Rename sexual system    
        mutate(sexual_system_modified = case_when(
            sexual_system == "B"  ~ "ssher")) %>%   
        
        add_column(source = "https://sura.ots.ac.cr/florula4/docs/lista_arboles_sindromes_OVR05.pdf")  %>% 
        arrange(genero, especie)

```

### Species with no traits
```{r}
reactable(anti_join(species_with_no_traits_13,data_vargas_traits , by = "name_submitted"))
```
[Cecropia peltata info](https://www.srs.fs.usda.gov/pubs/misc/ag_654/volume_2/cecropia/peltata.htm)

```{r}
# Manual input
manual_input  <-   
    tribble(
         ~spcode, ~spcode_4_3, ~accepted_species, ~genero, ~especie, ~dispersal_syndrome_modified, ~pollination_syndrome_modified, ~sexual_system_modified, ~source,
        "cecrpe", "cecrpel", "Cecropia peltata", "Cecropia", "peltata","dani","dwin","ssdio",
        "http://www.ecofog.gf/img/pdf/bibliographic_synthesis_ruth_tchana_thomas_monjoin_master_2_bioget_cecropia.pdf",
        "tabral", "tabealb", "Tabernaemontana alba", "Tabernaemontana","alba","","","","",
        "pescar","tabearb", "Tabernaemontana arborea", "Tabernaemontana","arborea","","","",""
    )    
```

# Morphospecies

```{r}
# Get morpho-species
morpho_species <- 
  data_species_full_list %>%
        
        # Get morphospecies
        filter((str_detect(especie, "^sp") & (str_length(especie) <= 4))) %>% 
        
        # Set first letter to uppercase
        mutate(genero = str_to_title(genero)) %>% 
        
        select(-c(familia))  %>% 
        group_by(genero) %>% 
    
        # Name morpho-species in a sequential manner
        mutate(especie = paste0("sp", row_number())) %>% 
        
        unite("accepted_species", c(genero, especie), sep = " ", remove = F) %>% 
        
        arrange(accepted_species)

```


# Raw Reproductive traits

```{r}
# Raw reproductive traits
raw_reproductive_traits <- 
   full_join(
       
        # Salgado data
         data_salgado_traits %>%
             select(-c(name_submitted, genero, especie,
                       dispersal_syndrome,pollination_syndrome,
                       sexual_system)),
         #Chazdon Data
         data_chazdon_traits %>%
             select(-c(name_submitted, genero, especie,
                       dispersal_syndrome,pollination_syndrome,
                       sexual_system)),
         by = c("accepted_species", "dispersal_syndrome_modified",
                     "pollination_syndrome_modified", "sexual_system_modified",
                     "source","spcode", "spcode_4_3")) %>%   
    
    # Zamora data
    full_join(
        data_zamora_traits %>%
            select(-c(name_submitted, genero, especie,
                      dispersal_syndrome,pollination_syndrome,
                      sexual_system)),
        by = c("accepted_species", "dispersal_syndrome_modified",
               "pollination_syndrome_modified", "sexual_system_modified",
               "source","spcode", "spcode_4_3"))  %>% 

        
    # Vargas data
    full_join(., 
              data_vargas_traits %>% 
                  select(-c(name_submitted, genero, especie, sexual_system)),
              by = c("accepted_species", "dispersal_syndrome_modified",
                     "pollination_syndrome_modified", "sexual_system_modified",
                     "source","spcode", "spcode_4_3")) %>% 
                    
    # Add morpho species
    #  Here column genero is generated
    full_join(.,
              morpho_species %>% 
                  select(-c(genero, especie)),
              by = c("accepted_species","spcode_4_3", "spcode"))  %>% 
    
    # Manual input
    full_join(., 
              manual_input %>% 
                  select(-c(genero, especie)), 
              by = c("accepted_species", "dispersal_syndrome_modified",
                     "pollination_syndrome_modified", "sexual_system_modified",
                     "source", "spcode", "spcode_4_3")) %>% 
        
    # Replace empty values with NA
    mutate_all(na_if, "") %>% 
    select(-genero) 
    
```

```{r}
raw_reproductive_traits <- 
    raw_reproductive_traits %>% 
    
        # Change D-P H-D value 
        mutate(sexual_system_modified = case_when(
            sexual_system_modified == "D-P" | sexual_system_modified == "H-D" ~ "ssdio",
            TRUE ~ sexual_system_modified)) %>% 
        arrange(accepted_species)   
```


# Raw Wood density

```{r}
# Join the names from TRNS with the morpho-species

wood_density_sp_list <- 
        # Get morphospecies
        morpho_species %>% 
    
        # Create column for joining the datasets
        unite(accepted_species, genero, especie, sep = " ") %>% 
        
            
        # Join morpho-species with species with a full name n = 256    
        full_join(.,species_list,by = "accepted_species" )  %>%  
        
        # Get column the accepted_species with morpho species n = 256
        select(accepted_species) %>% 
    
        # Create columns for the function getWoodDensity
        separate(accepted_species, c("genus", "specie"), sep = " ", remove = F) %>% 
        
        # This step is done for calculating wood density at the genus level for
        # the morpho-species. Without it Morpho-species will get wd a the species level        
        mutate(specie = case_when(
            str_detect(specie, "^sp") & str_length(specie) <= 4 ~ "NA",
            TRUE ~ specie)) %>% 
        arrange(genus, specie)
        
```

## Get wood density values

```{r cache = TRUE, wood_density_cache, echo = TRUE}
# https://search.r-project.org/CRAN/refmans/BIOMASS/html/getWoodDensity.html
wood_density <- 
    getWoodDensity(
          genus = wood_density_sp_list$genus,
          species = wood_density_sp_list$specie,
          region = c("CentralAmericaTrop", "SouthAmericaTrop"))
```

```{r}
raw_wood_density <- 
    wood_density %>% 
        group_by(genus, species) %>% 
        
        # Name morpho-species in a sequential manner
        mutate(species = if_else(species == "NA",paste0("sp", 
                                                        row_number()),species)) %>%
        
        # Arrange columns and clean names
        select(1:3,levelWD, everything()) %>% 
        clean_names() %>% 
        arrange(genus, species) %>% 
        unite(accepted_species, c("genus","species"), sep = " ") %>% 
        select(-family) 
        
        # Check that morpho species get a wd at the level of genus 
        # filter((str_detect(accepted_species, pattern = "sp"))) %>% 
        # arrange(level_wd)
```

```{r}
# Raw wood density
raw_wood_density <- 
    raw_wood_density %>% 
        arrange(accepted_species)
```


# Join reproductive traits, wood density, seed length and leaf P and N data

## Load leaf P and N data

```{r}
source("./scripts/script_effect_traits_data_cleaning.R")

data_effect_traits_n_p <-  
    
    data_effect_traits %>% 
        select(spcode,spcode_4_3,accepted_species, n_mgg_1, p_mgg_1) %>% 
        mutate(n_p_ratio = n_mgg_1/p_mgg_1)
    
```

## Load seed length data

```{r}
source("./scripts/script_species_seed_length.R")
data_seed_length <- 
    data_seed_length %>% 
    mutate(seedlength_cat = as.character(seedlength_cat)) %>% 
     # Rename levels in pollination column
        mutate(seedlength_cat = case_when(
            
            seedlength_cat  == "1 to <6 mm"  ~ "1_to_6mm",
            seedlength_cat  == "6 to <15 mm"  ~ "6_to_15mm",
            
            seedlength_cat  == "15 to <30 mm"  ~ "15_to_30mm",
            seedlength_cat  == "30 to <50 mm"  ~ "30_to_50mm",
            seedlength_cat  == "50 to <100 mm"  ~ "50_to_100mm",
            seedlength_cat  == ">100 mm"  ~ ">100mm",
            seedlength_cat  == "<1 mm"  ~ "<1mm",
            seedlength_cat  == "under 1 mm"  ~ "<1mm",
            TRUE ~ seedlength_cat)) 
```
## Load seed mass data

```{r}
# Data is a mess check the others first
```


## Check dataset
```{r}
reactable(inner_join(raw_reproductive_traits, raw_wood_density, 
                     by = "accepted_species"),
          width = "2750",
          height = "750",)
```

## Write full dataset
```{r}
# db with morpho species 
traits_db_255 <- 
    inner_join(raw_reproductive_traits, raw_wood_density, 
                                by = c("accepted_species"))  %>% 
    select(!c(source,level_wd, sd_wd, n_ind )) %>% 
    
    # Add seed length
    full_join(., data_seed_length, by = c("spcode","spcode_4_3")) %>% 
    
    # Add leaf N and P
    full_join(.,data_effect_traits_n_p, by = c("spcode","spcode_4_3", "accepted_species")) %>% 
    
    # Mutate characters as factor
    mutate(across(where(is.character), as.factor)) %>% 
    arrange(spcode_4_3) %>% 
    dplyr::na_if("NA")
```



```{r}
# db no morpho species
traits_db_190 <- 
    inner_join(raw_reproductive_traits, raw_wood_density, 
                                by = c("accepted_species"))  %>% 
    select(!c(source,level_wd, sd_wd, n_ind )) %>% 
    
    # Add seed length
    full_join(., data_seed_length, by = c("spcode","spcode_4_3")) %>% 
    
    # Add leaf N and P
    full_join(.,data_effect_traits_n_p, by = c("spcode","spcode_4_3", "accepted_species")) %>% 
    
    # Mutate characters as factor
    mutate(across(where(is.character), as.factor)) %>% 
    arrange(spcode_4_3) %>% 
    dplyr::na_if("NA") %>%
    
    # Remove morphospecies aka species names with number 
    filter(!(str_detect(accepted_species, "[:digit:]")))
```

# Data checks
```{r}
# How many species have no seed data
traits_db_190 %>% 
    filter(is.na(seedlength_mm) & is.na(seedlength_cat)) %>% 
    nrow()
```


```{r}
# Species with 4 traits or less
traits_db_190 %>% 
    column_to_rownames("accepted_species") %>% 
    select(3:ncol(.))  %>% 
    mutate(species_number_of_traits = rowSums(!is.na(.))) %>% 
    filter(species_number_of_traits <= 4)
```



```{r}
write.csv(traits_db_190,"./data/data_for_analisys/db_traits_190.csv")
```

