---
title: 'Trait based analysis Redundancy of reproductive and wood density traits'
author: "Erick Calderon-Morales"
date: '2022'
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
					  fig.width = 11, fig.height = 7)
```

```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      
                      # Save a pdf_copy?
                      #dev            = c( "png", "pdf"),
                      # Include code?
                      echo           = TRUE,                      
                      error          = FALSE,
                      fig.align      = "center",
                      # Path where figures are going to be store pdf single figures
                      fig.path       = paste0("./notebooks/figures_trait_based_analisys", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```

```{r}
library(dplyr)
library(janitor)
library(tibble)
library(tidyr)
```

```{r}
# Load data
raw_data_complete <- read.csv("./raw_data/raw_data_complete.csv", header = T)
```

```{r}
# Remove columns that are not useful for the analysis
data_cleaned <-     
    raw_data_complete %>% 
        select(-c(level_wd,sd_wd,n_ind_wd, X, source)) %>% 
        column_to_rownames("accepted_species") %>% 
        
        # Rename columns
        rename(disp_sym = "dispersal_syndrome_modified",
               poli_sym = "pollination_syndrome_modified",
               ss_syst = "sexual_system_modified",
               wd = "mean_wd") %>% 
        mutate(across(where(is.character), factor))
        
        # Transform catergorical traits to factor 


```

# Chapter 3: Trait dissimilarity

```{r libaries, message=FALSE, warning=FALSE, cache=FALSE}
library(FD)
library(gtools)
library(gawdis)
library(NbClust)
library(gawdis)
```

```{r}
# Load trova function
source(here::here("../../cursos_libros_tutoriales/libros/handbook_of_trait_based_ecology/chapters_code/chapter3/trova.r"))
```

## 3.3 Calculation of trait dissimilarity

NA


## 3.4 Problems with the Gower distance

If we compute the dissimilarity for each trait and them combine these two 
dissimilarities with a simple average (or even with Euclidean distance, 
i.e. geometric mean), the resulting dissimilarity will be affected much more by 
the binary trait than by the quantitative. In other words, the contribution of 
the binary trait to the overall dissimilarity will be disproportional.


## 3.5 Categorical and fuzzy coded traits

```{r}
data_cleaned_dummy <- 
    cbind(acm.disjonctif(data_cleaned[1:3]),data_cleaned$wd) %>%
        clean_names()

# Trait dispersal syndrome
#rowSums(data_cleaned_dummy[1:4])
```

## 3.7 Functional groups with students’ traits

```{r}
par(mfrow = c(2, 2))
hist(data_cleaned$wd, col = "grey", main = "wood density trait values")
hist(log(data_cleaned$wd), col = "grey", main = "wood density trait values trait values log")
hist(dist(data_cleaned$wd), col = "grey", main = "dissimilarities wd")
hist(dist(log(data_cleaned$wd)), col = "grey", main = "dissimilarities wd log")
```


```{r, distance_matrix, cache = TRUE}

distance_matrix <- gawdis(data_cleaned, 
                    w.type = "optimized",
                    opti.maxiter = 300)

distance_matrix <- round(distance_matrix,3)
```

```{r}
pcoa_all <- dudi.pco(sqrt(distance_matrix), scannf = F, nf = 10)
scatter(pcoa_all)
```



```{r}
# We can compute how much of the variability is captured by the first two axes, 
# which is around 45%.
sum(pcoa_all$eig[1:2]) / sum(pcoa_all$eig)
```

```{r}
cluster <- hclust(distance_matrix, method = "ward.D2")
```

We can try one of many possible ways to find, statistically, the optimal 
number of groups in such a cluster (which maximizes the dissimilarity between 
groups and minimize the dissimilarity within groups). A good function is 
NbClust which offers a lot of possible approaches:


```{r}
cluster_groups <- NbClust(diss = distance_matrix, distance = NULL, min.nc = 5,
                      method = "ward.D", index = "silhouette")

```

```{r}
cluster_groups$Best.nc
```

```{r fig.height = 15, fig.width = 15}
library(factoextra)
library(ggsci)
fviz_dend(cluster,
          k = cluster_groups$Best.nc[1],
          k_colors = "jco",
          repel = TRUE,
          horiz = TRUE,
          type = "circular")
```



We can also visualize the groups using s.class, with respect to the PCoA:
```{r}
par(mfrow = c(1, 2))
scatter(pcoa_all)
s.class(pcoa_all$li, as.factor(cluster_groups$Best.partition), cpoint = 1)
```

## 3.8 Some real data with the function trova

The function allows to compute species dissimilarity not only using species 
means, but also considering overlap of trait values between species by using 
the standard deviation of a give trait.

```{r, eval = FALSE}
data_cleaned_dummy <- data_cleaned_dummy[, c(1:4,14)]
list_dummy <- list(data_cleaned_dummy[, 1:4])
names(list_dummy) <- "disp_symd"

dissims <- trova(species.names = rownames(data_cleaned_dummy), 
                 # needs mean and sd
                 gaussian.data = wd,
                 multiple.category = list_dummy)
```

## 3.9 The gawdis function

The dissimilarity obtained can be computed in order to attain a quasi-identical 
contribution of individual variables (e.g. traits) or group of associated 
variables (on variables reflecting similar information, e.g. multiple leaf traits)

```{r}
dummy$trait
```

```{r eval = FALSE}
equalcont <- gawdis(data_cleaned, 
                    w.type = "optimized",
                    opti.maxiter = 300)

equalcont <- round(equalcont,3)
attr(equalcont, "correls")
attr(equalcont, "weights")

```


## The function gawdis: grouping traits

```{r}
# load shade house experiment
load("../../data_test_shadehouse.RData")

traits <- 
    data_complete %>%
        unite("group", 1:4, sep = "_") %>% 
        column_to_rownames("group") %>%  
        select(8:16,20,21)  
    
```

```{r}
straightgowdis_2 <- gawdis(traits, w.type = "equal", silent = T)
cors_gow <- attr(straightgowdis_2, "correls")
```

```{r eval = FALSE}
gaw_groups <- gawdis(traits, 
                     w.type = "optimized", 
                     opti.maxiter = 300,
                     groups.weight = T, 
                     groups = c(1, 1, 1, 2, 2, 2, 2, 3, 3, 2))
```
```{r eval = FALSE}
cors_gaw_gr <- attr(gaw_groups, "correls")
ncol(traits)

cors_gaw_gr[11] <- attr(gaw_groups, "group.correls")[1]
names(cors_gaw_gr)[11] <- "phys"

cors_gaw_gr[12] <- attr(gaw_groups, "group.correls")[2]
names(cors_gaw_gr)[12] <- "leaf_traits"

cors_gaw_gr[13] <- attr(gaw_groups, "group.correls")[3]
names(cors_gaw_gr)[13] <- "isotopes"

```
# Chapter 4: (Multivariate) species level responses

```{r}
data_path <- "../../cursos_libros_tutoriales/libros/handbook_of_trait_based_ecology/chapters_code/chapter4/"

```


```{r}
com <- read.delim(file = paste0(data_path,"Site_species.txt"))
community <- read.csv("")
```

```{r}
traits <- read.delim(file = here::here("data", "chapter4", "Species_traits.txt"))
```

```{r}
env <- read.delim(here::here("data", "chapter4", "Site_env.txt"))
```

