---
title: 'CWM'
author: "Erick Calderon-Morales"
date: '2022'
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
					  fig.width = 11, fig.height = 7)
```

```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      
                      # Save a pdf_copy?
                      #dev            = c( "png", "pdf"),
                      # Include code?
                      echo           = TRUE,                      
                      error          = FALSE,
                      fig.align      = "center",
                      # Path where figures are going to be store pdf single figures 
                      fig.path       = paste0("./notebooks/figures_cwm", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```



```{r}
library(FD)
library(psych)
library(tidyr)
library(dplyr)
library(picante)
library(gawdis)
```

```{r}
#source(here::here("./codes/scripts/script_agb_data_cleaning.R"))
source(here::here("./codes/scripts/script_abundance_data_cleaning.R"))
source(here::here("./codes/scripts/script_env_data_cleaning.R"))
```


```{r}
# Load effect traits
source(here::here("./codes/scripts/script_effect_traits_data_cleaning.R"))

traits_effect <-  
        data_effect_traits %>% 
        dplyr::select(-c(spcode,accepted_species)) 
        #arrange(spcode_4_3)  
        #column_to_rownames("spcode_4_3")

```

```{r}
traits_response <- 
    read.csv("./data_for_analisys/response_traits.csv", header = T) %>% 
        
        dplyr::select(3,5:7,9) %>% 
        #column_to_rownames("spcode_4_3") %>% 
        mutate(across(where(is.character), as.factor))
    
```


```{r}
traits <- 
    inner_join(traits_effect, traits_response, by = "spcode_4_3") %>% 
        select(-c(af_mm2, afe_mm2mg_1,n_mgg_1, dm_gcm3_1,cfms_mgg_1))  %>% 
        column_to_rownames("spcode_4_3")
```



```{r}
community <- 
    data_abundance_new_spcodes %>% 
    pivot_longer(!parcela,names_to = "spcode", values_to = "abundance") %>% 
    pivot_wider(names_from = parcela, values_from = abundance) %>% 
    arrange(spcode) %>% 
    column_to_rownames("spcode")
```

```{r}
env <- 
    data_env %>% 
        column_to_rownames("plot") %>% 
        select(-forest_type)
```

```{r}
env_cor <- cor(env)
env_cor[,"elev"]

```

```{r}

```


# The MAJORITY of species in low-elevation habitats have given traits
```{r}

```



# Functional diversity
```{r}
dim(traits_effect)
dim(community)

#fd_effect <- dbFD(traits_effect, t(community), CWM.type = "all")
```


```{r}
dist_traits_response <- gawdis(traits, 
                    w.type = "optimized",
                    opti.maxiter = 400)
```


```{r}
#w <- c(0.3167144,0.3496200,0.3336655)
fd_response <- dbFD(dist_traits_response, t(community),corr = "cailliez")

```


```{r}
indices <- cbind(fd_response$nbsp, fd_response$FRic, fd_response$FEve, 
                 fd_response$FDiv, fd_response$FDis, fd_response$RaoQ)
colnames(indices) <- c("NumbSpecies", "FRic", "FEve", "FDiv", "FDis", "Rao")
pairs(indices, pch = 20)
```

```{r}
par(mfrow = c(2, 2)) #this is just to prepare the space
par(mar = c(4, 4, 2, 1)) # margin of the plots, just to

plot(env$elev, fd_response$nbsp, xlab = "moisture",
     ylab = "Number of species", pch = 20)

plot(env$elev, fd_response$FRic, xlab = "moisture",
     ylab = "Functional Richness", pch = 20)

plot(env$elev, fd_response$FEve, xlab = "moisture", 
        ylab = "Functional Evenness", pch = 20)

plot(env$elev, fd_response$RaoQ, xlab = "moisture",
     ylab = "Functional Dispersion", pch = 20)
```

```{r}
summary(lm(fd_response$FEve ~ preccv , data = env))
```

```{r}
source(here::here("../../cursos_libros_tutoriales/libros/handbook_of_trait_based_ecology/chapters_code/chapter5/Rao.r"))
```

```{r}
test <- Rao(log(community + 1), dist_traits_response, dphyl = NULL, 
            weight = F, Jost = T, structure = NULL)
```

```{r}
test$TD$Beta_prop
```


We can see in this figure that, despite the great turnover in species composition 
(beta TD, grey colour, left stack) we have a real low functional turnover. In 
other words most of the trait dissimilarity between species is found within a 
plot and not across plots, even if there are such a marked environmental changes 
and even if species composition changes are very strong (high beta TD)

```{r}
TD.NEspain.prop <- c(100 * test$TD$Mean_Alpha / test$TD$Gamma, test$TD$Beta_prop)

FD.NEspain.prop <- c(100* test$FD$Mean_Alpha/test$FD$Gamma, test$FD$Beta_prop)

barplot(cbind(TD.NEspain.prop, FD.NEspain.prop), beside = F, col = c(8, 0),
ylim = c(0, 100), ylab = c("Proportion of Alpha and Beta in Equivalent numbers"))
text(0.7, 60, "beta")
text(0.7, 15, "alpha")
```

```{r}

```

