---
title: 'Functinanl diversity measures'
author: "Erick Calderon-Morales"
date: '2022'
output:
  prettydoc::html_pretty:
    highlight: pygments
    theme: cayman
    toc: yes
    number_sections: no
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,comment = "", fig.align = 'center',
					  fig.width = 11, fig.height = 7)
```

```{r knitr, include = FALSE}

# Save figures in specific place

knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.comments = TRUE,
                      
                      # Save a pdf_copy?
                      #dev            = c( "png", "pdf"),
                      # Include code?
                      echo           = TRUE,                      
                      error          = FALSE,
                      fig.align      = "center",
                      # Path where figures are going to be store pdf single figures 
                      fig.path       = paste0("./notebooks/figures_trait_based_analisys", "/"),
                      fig.width      = 11,
                      fig.height     = 7,
                      message        = FALSE,
                      warning        = FALSE)
```


```{r}
library(dplyr)
library(janitor)
library(tibble)
library(tidyr)
```

```{r}
# Load data
raw_data_complete <- read.csv("./data_for_analisys/traits_data.csv", header = T)
```

```{r}
# Remove columns that are not useful for the analysis
data_cleaned <-     
    raw_data_complete %>% 
        select(-c(level_wd,sd_wd,n_ind_wd, X, source, spcode, spcode, spcode_4_3)) %>% 
        column_to_rownames("accepted_species") %>% 
        
        # Rename columns
        rename(disp_sym = "dispersal_syndrome_modified",
               poli_sym = "pollination_syndrome_modified",
               ss_syst = "sexual_system_modified",
               wd = "mean_wd") %>% 
        mutate(across(where(is.character), factor))
        
        # Transform catergorical traits to factor 


```

# Chapter 3: Trait dissimilarity

```{r libaries, message=FALSE, warning=FALSE, cache=FALSE}
library(FD)
library(gtools)
library(gawdis)
library(NbClust)
library(gawdis)
```


```{r}
# Load trova function
source(here::here("../../cursos_libros_tutoriales/libros/handbook_of_trait_based_ecology/chapters_code/chapter3/trova.r"))
```

## 3.3 Calculation of trait dissimilarity

NA


## 3.4 Problems with the Gower distance

If we compute the dissimilarity for each trait and them combine these two 
dissimilarities with a simple average (or even with Euclidean distance, 
i.e. geometric mean), the resulting dissimilarity will be affected much more by 
the binary trait than by the quantitative. In other words, the contribution of 
the binary trait to the overall dissimilarity will be disproportional.


## 3.5 Categorical and fuzzy coded traits

```{r}
data_cleaned_dummy <- 
    cbind(acm.disjonctif(data_cleaned[1:3]),data_cleaned$wd) %>%
        clean_names()

# Trait dispersal syndrome
rowSums(data_cleaned_dummy[1:4])
```

## 3.7 Functional groups with students’ traits

```{r}
par(mfrow = c(2, 2))
hist(data_cleaned$wd, col = "grey", main = "wood density trait values")
hist(log(data_cleaned$wd), col = "grey", main = "wood density trait values trait values log")
hist(dist(data_cleaned$wd), col = "grey", main = "dissimilarities wd")
hist(dist(log(data_cleaned$wd)), col = "grey", main = "dissimilarities wd log")
```


```{r, distance_matrix, cache = TRUE}

distance_matrix <- gawdis(data_cleaned, 
                    w.type = "optimized",
                    opti.maxiter = 500)

distance_matrix <- round(distance_matrix,3)
attr(distance_matrix, "correls")
attr(distance_matrix, "weights")
```

```{r}
pcoa_all <- dudi.pco(sqrt(distance_matrix), scannf = F, nf = 10)
scatter(pcoa_all)
```



```{r}
# We can compute how much of the variability is captured by the first two axes, 
# which is around 45%.
sum(pcoa_all$eig[1:2]) / sum(pcoa_all$eig)
```

```{r}
cluster <- hclust(distance_matrix, method = "ward.D2")
```

We can try one of many possible ways to find, statistically, the optimal 
number of groups in such a cluster (which maximizes the dissimilarity between 
groups and minimize the dissimilarity within groups). A good function is 
NbClust which offers a lot of possible approaches:


```{r}
cluster_groups <- NbClust(diss = distance_matrix, distance = NULL, min.nc = 6,
                      method = "ward.D", index = "silhouette")
```

```{r}
cluster_groups$Best.nc
```

```{r fig.height = 20, fig.width = 20}
library(factoextra)
library(ggsci)
fviz_dend(cluster,
          k = cluster_groups$Best.nc[1],
          k_colors = "jco",
          repel = TRUE,
          horiz = TRUE,
          type = "circular")
```



We can also visualize the groups using s.class, with respect to the PCoA:
```{r}
par(mfrow = c(1, 2))
scatter(pcoa_all)
s.class(pcoa_all$li, as.factor(cluster_groups$Best.partition), cpoint = 1)
```

## 3.8 Some real data with the function trova

The function allows to compute species dissimilarity not only using species 
means, but also considering overlap of trait values between species by using 
the standard deviation of a give trait.

```{r, eval = FALSE}
data_cleaned_dummy <- data_cleaned_dummy[, c(1:4,14)]
list_dummy <- list(data_cleaned_dummy[, 1:4])
names(list_dummy) <- "disp_symd"

dissims <- trova(species.names = rownames(data_cleaned_dummy), 
                 # needs mean and sd
                 gaussian.data = wd,
                 multiple.category = list_dummy)
```

## 3.9 The gawdis function

The dissimilarity obtained can be computed in order to attain a quasi-identical 
contribution of individual variables (e.g. traits) or group of associated 
variables (on variables reflecting similar information, e.g. multiple leaf traits)

```{r}
dummy$trait
```

```{r eval = FALSE}
equalcont <- gawdis(data_cleaned, 
                    w.type = "optimized",
                    opti.maxiter = 300)
```


## The function gawdis: grouping traits

```{r}
# load shade house experiment
load("../../data_test_shadehouse.RData")

traits <- 
    data_complete %>%
        unite("group", 1:4, sep = "_") %>% 
        column_to_rownames("group") %>%  
        select(8:16,20,21)  
    
```

```{r}
straightgowdis_2 <- gawdis(traits, w.type = "equal", silent = T)
cors_gow <- attr(straightgowdis_2, "correls")
```

```{r eval = FALSE}
gaw_groups <- gawdis(traits, 
                     w.type = "optimized", 
                     opti.maxiter = 300,
                     groups.weight = T, 
                     groups = c(1, 1, 1, 2, 2, 2, 2, 3, 3, 2))
```

```{r eval = FALSE}
cors_gaw_gr <- attr(gaw_groups, "correls")
ncol(traits)

cors_gaw_gr[11] <- attr(gaw_groups, "group.correls")[1]
names(cors_gaw_gr)[11] <- "phys"

cors_gaw_gr[12] <- attr(gaw_groups, "group.correls")[2]
names(cors_gaw_gr)[12] <- "leaf_traits"

cors_gaw_gr[13] <- attr(gaw_groups, "group.correls")[3]
names(cors_gaw_gr)[13] <- "isotopes"
```

# Chapter 4: (Multivariate) species level responses


```{r}
# Packages 
library(ade4)
library(MASS)
library(vegan)
library(ecodist)
library(maptools)
library(rpart)
library(splines)
library(gam)
library(pgirmess)
library(utils)
library(combinat)
library(cluster)
library(fpc)
library(clusterSim)
library(lmtest)
library(Hmisc)
library(gplots)
library(NbClust)
library(rpart)
library(rpart.plot)
library(dismo)
library(multcomp)
library(gbm)
library(raster)


#source(here::here("./codes/scripts/script_agb_data_cleaning.R"))
source(here::here("./codes/scripts/script_abundance_data_cleaning.R"))
source(here::here("./codes/scripts/script_env_data_cleaning.R"))
```

```{r}
data_path <- "../../cursos_libros_tutoriales/libros/handbook_of_trait_based_ecology/chapters_code/chapter4/"

# Additional functions
source(here::here(data_path,"Inference_modelset.r"))
source(here::here(data_path,"Inference_compute.r"))
source(here::here(data_path,"corratio.R"))
source(here::here(data_path,"calinski.R"))
source(here::here(data_path,"VarScoreOMI.r"))
source(here::here(data_path,"doublerda.R"))
```





```{r}
comm <- read.delim(file = paste0(data_path,"Site_species.txt"))
my_community <- 
    data_abundance_new_spcodes %>% 
    column_to_rownames("parcela")


```

```{r}
traits <- read.delim(file = here::here(data_path,"Species_traits.txt"))

my_traits <- 
    read.csv("./data_for_analisys/traits_data.csv", header = T) %>% 
        dplyr::select(3,5:7,9) %>% 
        column_to_rownames("spcode_4_3") %>% 
        mutate(across(where(is.character), as.factor))
    
```

```{r}
env <- read.delim(here::here(data_path, "Site_env.txt"))

my_env <- 
    data_env %>% 
        column_to_rownames("plot") %>% 
        dplyr::select(-c(crtm_90_x,crtm_90_y,forest_type))
    
```

## “Trait-free” Canonical Correspondance Analysis
 
```{r}
cca1 <- cca(my_community ~ sand + limo + clay + p_h + acidity +
                ca + mg + k + p + organic_matter +
                elev + slope_per + slope_deg +
                prec + precdriest + preccv + temp + tempmin, data = my_env
            )
anova(cca1, by = "terms")
RsquareAdj(cca1)
```

```{r}
plot(cca1, display = c("species", "bp"))
```

```{r}
# we might be interested, is the position for each species on the axes of 
# the CCA ordination space
head(vegan::scores(cca1, display = "species"))
```
```{r}

cca_preccv <- cca(my_community ~ preccv + Condition(sand + limo + clay + p_h + 
                                                    acidity + ca + mg + k + p + 
                                                    organic_matter + elev + 
                                                    slope_per + slope_deg +
                                                    prec + preccv + temp +
                                                    tempmin), 
                data = my_env)

cca_p <- cca(my_community ~ p + Condition(sand + limo + clay + p_h + 
                                                    acidity + ca + mg + k +  
                                                    organic_matter + elev + 
                                                    slope_per + slope_deg +
                                                    prec + preccv + temp +
                                                    tempmin), 
            data = my_env)

cca_organic_matter <- cca(my_community ~ organic_matter + 
                              Condition(sand + limo + clay + p_h + acidity + 
                                            ca + mg + k + elev + p + 
                                            slope_per + slope_deg + 
                                            prec + preccv + temp + tempmin), 
            data = my_env)
```

```{r}
plot(cca_p, display = c("species", "bp"))
plot(cca_preccv, display = c("species", "bp"))
```

## Double Canonical Correspondance Analysis

```{r}
ca1_book <- dudi.coa(comm, scannf = F)
dCCA1_book <- dbrda(ca1_book, env, traits, scannf = FALSE)

```

```{r}
#dummy_traits <-  
#    cbind(acm.disjonctif(my_traits[1:3]), my_traits$mean_wd) %>%
#        clean_names()


ca1 <- dudi.coa(my_community, scannf = F)

dCCA1 <- dbrda(ca1, my_env, data_effect_traits_cleaned, scannf = FALSE)
```

```{r}
sum(dCCA1$eig) / sum(ca1$eig)
```

```{r}
dCCA1$eig / sum(dCCA1$eig)
```

```{r fig.width = 10, fig.width = 10}
s.arrow(dCCA1$corZ[-1, ], ylim = c(-1.2, 1.2), boxes = FALSE)
s.arrow(dCCA1$corX[-1, ], add.plot = T, clab = 1.5)
pointLabel(dCCA1$co, row.names(dCCA1$co), cex = 0.5, col = "blue")
```

## Functional response groups

```{r}
# Load effect traits
source(here::here("./codes/scripts/script_effect_traits_data_cleaning.R"))
```


```{r}
dist_CCA <- dist(vegan::scores(cca1, display = "species"))
clust_CCA <- hclust(dist_CCA, method = "ward.D2")
plot(clust_CCA, cex = 0.6)
```

```{r}
groups_CCA <- NbClust(diss = dist_CCA, distance = NULL, min.nc = 3, 
                      method = "ward.D2", index = "silhouette")
```

```{r}
# Load effect traits
source(here::here("./codes/scripts/script_effect_traits_data_cleaning.R"))
data_effect_traits_cleaned <-  
        data_effect_traits %>% 
        dplyr::select(-c(spcode,accepted_species)) %>% 
        column_to_rownames("spcode_4_3")

summary(lm(as.matrix(data_effect_traits_cleaned) ~ groups_CCA$Best.partition))
```
```{r}
dist_dCCA <- dist(dCCA1$co)
clust_dCCA <- hclust(dist_dCCA, method = "ward.D2")
plot(clust_dCCA, cex = 0.6)
```

```{r}
groups_dCCA <- NbClust(diss = dist_dCCA, distance = NULL, min.nc = 2, max.nc = 9, 
                       method = "ward.D2", index = "silhouette")
```
```{r fig.width= 10, fig.height=10}
clusters <- as.factor(groups_dCCA$Best.partition)
s.class(dCCA1$co, clusters, col = 1:nlevels(clusters))
s.arrow(dCCA1$corZ[-1,], add.plot = T, clab = 0.8, boxes = FALSE)
s.arrow(dCCA1$corX[-1, ], add.plot = T, clab = 1.2)
```

## RDA and regression trees

```{r}
rda1 <- rda(my_community ~ ., data = my_env, scale = TRUE)
plot(rda1, display = c("bp", "sp"))
```

```{r}
RsquareAdj(rda1)
```


```{r}
cumsum(rda1$CCA$eig) / sum(rda1$CCA$eig)
```
```{r}
lm_rda_effect <- lm(vegan::scores(rda1, choices = 1, display = "species") ~ .
                    , data = data_effect_traits_cleaned)
summary(lm_rda_effect)
```


## Regression Trees

```{r}
# Create a data frame that contains both, the response (i.e. species scores from
# RDA axes 1 and 2) and the traits to model these responses.

scores_rda1 <- vegan::scores(rda1, choices = c(1:12))$species

df_effect <- cbind(scores_rda1, data_effect_traits_cleaned)
df_response <- cbind(scores_rda1, my_traits)


head(df_response)
```
```{r}
rta <- rpart::rpart(RDA2 ~ af_mm2 + afe_mm2mg_1 + cfms_mgg_1 + dm_gcm3_1 + 
                        n_mgg_1 + p_mgg_1, data = df_effect)


rta_response <- rpart::rpart(RDA5 ~ dispersal_syndrome_modified + 
                        pollination_syndrome_modified + 
                        sexual_system_modified , data = df_response)

rpart.plot(rta_response)
```


```{r}
abundance_long <- 
    my_community %>% 
        rownames_to_column("plot") %>% 
        pivot_longer(!plot,names_to = "spcode", values_to = "abundance") %>% 
        group_by(spcode) %>% 
        # Relative abundance
        summarise(abundance = sum(abundance)) %>% 
        mutate(relative_abundance = abundance/sum(abundance)) 
    
```


```{r}
#env_log <- 
    my_env 
        rownames_to_column("spcode") %>% 
        full_join(., abundance_long, by = "spcode") %>%   
        dplyr::select(spcode, abundance, relative_abundance, everything()) %>%  
        mutate(across(where(is.numeric), log)) %>% 
        column_to_rownames("spcode") %>% 
        dplyr::select(-af_mm2)
```

```{r}
treeVegetFull <- rpart(abundance ~ af_mm2 + afe_mm2mg_1 + cfms_mgg_1 + dm_gcm3_1, 
                       n_mgg_1 + p_mgg_1,data = fitness_log,
                       control = rpart.control(maxdepth = 5))
rpart.plot(treeVegetFull)
```


## A “homemade” bagging example

```{r}
nreps <- 5000
#Create a matrix to store the predictions of each tree:
predictions <- matrix(NA, nrow = nrow(fitness_log), ncol = nreps,
                      dimnames = list(rownames(fitness_log), 
                                      paste("Tree", 1:nreps, sep = "."))
)
```


```{r}
for (i in 1:nreps) { #For each repetition
    
    #Observations in training set by means of bootstrap:
    selected <- sample(1:nrow(fitness_log), replace = TRUE)
    
    ##Note: selected has the same number of observation than meadows, but some are
    # repeated and some do not appear (about 1/3). You can check by
    # length(unique(selected)) #Approx. 2/3 of the observations
    OOBAux <- fitness_log[-selected, ]
    trainingAux <- fitness_log[selected, ]
    #Lets fit our tree, using trainingAux:
    
    treeAux <- rpart(relative_abundance ~ afe_mm2mg_1 + cfms_mgg_1 + dm_gcm3_1 +
                                n_mgg_1 + p_mgg_1, data = trainingAux,
                     control = rpart.control(cp = 0.001))
    #Lets predict for the OOB, and store the predictions in the matrix we created before the lo
    predictions[-selected, i] <- predict(treeAux, newdata = OOBAux)
    if (i %% 500 == 0) {
        cat(paste("All is good! Rep. number", i, "\n"))
    }
}
```

```{r}
Result <- rowMeans(predictions, na.rm = T)
```


```{r}
#And let us examine how much they resemble the real observations:
corBagging <- cor(fitness_log$relative_abundance, Result)
plot(fitness_log$relative_abundance ~ Result, main = paste("Bagging cor=", round(corBagging, 3)))
```

```{r}
MSE <- rep(NA, nreps)
for (i in 20:nreps) {
    #Let us start in 20 instead of 1, to be more or less sure that all observations
    #have been part of the OOB sample at least once
    averageAux <- rowMeans(predictions[, 1:i], na.rm = T)
    MSE[i] <- mean((fitness_log$relative_abundance - averageAux) ^ 2)
}
plot(20:nreps, MSE[20:nreps],type = "l")
```



```{r}
nProcesses <- 10
nreps <- 3000
plot(0, 0, type = "n", xlim = c(0, nreps), ylim = c(0, 1), ylab = "MSE",
     xlab = "N TREES")

for (k in 1:nProcesses) {
    predictions <- matrix(NA, nrow = nrow(fitness_log), ncol = nreps,
                          dimnames = list(rownames(fitness_log),
                                          paste("A", 1:nreps, sep = ".")))
    for (i in 1:nreps) {
        selected <- sample(1:nrow(fitness_log), replace = TRUE)
        OOBAux <- fitness_log[-selected, ]
        trainingAux <- fitness_log[selected, ]
        treeAux <- rpart(relative_abundance ~  afe_mm2mg_1 + cfms_mgg_1 + dm_gcm3_1 +
                         n_mgg_1 + p_mgg_1, data = trainingAux,
                         control = rpart.control(cp = 0.001))
        predictions[-selected, i] <- predict(treeAux, newdata = OOBAux)
    }
    MSE <- rep(NA, nreps)
    for (i in 20:nreps) {
        averageAux <- rowMeans(predictions[, 1:i], na.rm = T)
        MSE[i] <- mean((fitness_log$relative_abundance - averageAux) ^ 2)
    }
    lines(20:nreps, MSE[20:nreps], col = k)
}
```


## Boosted regression trees for studying the relationship between traits and abundance

```{r}
BRT.meadows.LR0.1 <- gbm.step(data = fitness_log,
                              gbm.x = c(3:7), #columns of predictors
                              gbm.y = 2, # Column of the response variable
                              family = "gaussian",
                              tree.complexity = 1,
                              learning.rate = 0.1,
                              bag.fraction = 0.75,
                              n.trees = 50,
                              verbose = FALSE)

```

```{r}
BRT.meadows.LR0.01 <- gbm.step(data = fitness_log,
                               gbm.x = c(3:7), #columns of predictors
                               gbm.y = 2, # Column of the response variable
                               family = "gaussian",
                               tree.complexity = 1,
                               learning.rate = 0.01,
                               bag.fraction = 0.75,
                               n.trees = 50,
                               verbose = FALSE)
```

```{r}
BRT.meadows.LR0.001 <- gbm.step(data = fitness_log,
                                gbm.x = c(3:7), #columns of predictors
                                gbm.y = 2, # Column of the response variable
                                family = "gaussian",
                                tree.complexity = 1,
                                learning.rate = 0.001,
                                bag.fraction = 0.75,
                                n.trees = 50,
                                verbose = FALSE)
```

```{r}
predicted <- predict(BRT.meadows.LR0.001, n.trees = BRT.meadows.LR0.001$n.trees)
observed <- fitness_log$abundance
plot(predicted ~ observed)
abline(0, 1, col = 2)
```

```{r}
R2BRT.meadows <- 1 - (BRT.meadows.LR0.001$self.statistics$mean.resid / 
                          BRT.meadows.LR0.001$self.statistics$mean.null)
R2BRT.meadows
```

```{r}
summary(BRT.meadows.LR0.001)
```
```{r}
gbm.interactions(BRT.meadows.LR0.001)$rank.list
```


Traits do not predict relative abundance

# Chapter 5

```{r}
library(FD)
library(psych)
library(tidyr)
library(dplyr)
```

```{r}
#source(here::here("./codes/scripts/script_agb_data_cleaning.R"))
source(here::here("./codes/scripts/script_abundance_data_cleaning.R"))
source(here::here("./codes/scripts/script_env_data_cleaning.R"))
```

```{r}
my_traits_response <- 
    read.csv("./data_for_analisys/response_traits.csv", header = T) %>% 
        dplyr::select(3,5:7,9) %>% 
        column_to_rownames("spcode_4_3") %>% 
        mutate(across(where(is.character), as.factor))
```


```{r}
# Load effect traits
source(here::here("./codes/scripts/script_effect_traits_data_cleaning.R"))
my_traits_effect <-  
        data_effect_traits %>% 
        dplyr::select(-c(spcode,accepted_species)) %>% 
        column_to_rownames("spcode_4_3")

```

```{r}
my_community <- 
    data_abundance_new_spcodes %>% 
    pivot_longer(!parcela,names_to = "spcode", values_to = "abundance") %>% 
    pivot_wider(names_from = parcela, values_from = abundance)    %>% 
    column_to_rownames("spcode")
```



```{r}
my_community_relative_abun <- 
    my_community %>% 
        rownames_to_column("spcode") %>% 
        pivot_longer(!spcode,names_to = "parcela", values_to = "abundance") %>% 
        mutate(across(where(is.character), as.factor)) %>% 
        group_by(parcela,spcode) %>% 
        tally(abundance) %>% 
        mutate(relative_abund = round(n/(sum(n)),15)) %>%  
        arrange(parcela)  %>% 
        select(-n)  %>% 
        pivot_wider(names_from = parcela, values_from = relative_abund) %>% 
        column_to_rownames("spcode")
```



```{r}
Bodysize <- c(10, 20, 30, 40, 50, NA, 70)
Carnivory <- c(1, 1, 0, 1, 0, 1, 0)
spxtraits.matrix <- cbind(Bodysize, Carnivory)
rownames(spxtraits.matrix) <- paste("species", 1:7)
spxtraits.matrix
```
```{r}
comm1 <- c(10, 10, 10, 10, 10, 0, 0)
comm2 <- c(49, 0, 5, 0, 5, 0, 11)
comm3 <- c(0, 6, 0, 4, 0, 2, 8)
spxcom.matrix <- cbind(comm1, comm2, comm3)
rownames(spxcom.matrix) <- paste("species", 1:7)
spxcom.matrix
```


```{r}
# Test 1
sp.rel.abb <- t(spxcom.matrix) / colSums(spxcom.matrix)
colSums(t(sp.rel.abb))
```


```{r}
# Test 2
sp.rel.abb2 <- round(t(my_community) / colSums(my_community),15)
colSums(t(sp.rel.abb2))
```


```{r}
# Test 3
sp.rel.abb3 <- my_community_relative_abun
colSums(sp.rel.abb3)

```

```{r}
t(sp.rel.abb) * spxtraits.matrix[, "Carnivory"]
colSums(t(sp.rel.abb) * spxtraits.matrix[, "Carnivory"])
```

```{r}
t(sp.rel.abb2) * my_traits_effect[, "afe_mm2mg_1"]
round(colSums(t(sp.rel.abb2) * my_traits_effect[, "afe_mm2mg_1"]),3)
```


```{r}
t(sp.rel.abb3) * my_traits_effect[, "afe_mm2mg_1"]
round(colSums(sp.rel.abb3 * my_traits_effect[, "afe_mm2mg_1"]),3)
```

## FD

```{r}
rownames(spxtraits.matrix) == rownames(spxcom.matrix)
functcomp(spxtraits.matrix, t(spxcom.matrix),CWM.type = "all")
```


```{r}
rownames(my_traits_effect) == rownames(my_community)
functcomp(my_traits_effect, t(my_community),CWM.type = "all")
```

## Calculation of CWM with real data

```{r}
my_traits_effect
my_traits_response
```

```{r}
rownames(my_community) == rownames(my_traits_effect)
rownames(my_community) == rownames(my_traits_response)
rownames(my_traits_effect) == rownames(my_traits_response)
```
```{r}
par(mfrow = c(1, 2))
par(mar = c(2, 4, 2, 0.5))
hist(my_traits_effect$afe_mm2mg_1 , main = "SLA", xlab = "")
hist(log(my_traits_effect$afe_mm2mg_1), main = "log(SLA)", xlab = "")
```
```{r}
par(mfrow = c(1, 2))
par(mar = c(2, 4, 2, 0.5))
hist(my_traits_effect$cfms_mgg_1 , main = "SLA", xlab = "")
hist(log(my_traits_effect$cfms_mgg_1), main = "log(SLA)", xlab = "")
```
```{r}
par(mfrow = c(1, 2))
par(mar = c(2, 4, 2, 0.5))
hist(my_traits_effect$dm_gcm3_1 , main = "SLA", xlab = "")
hist(log(my_traits_effect$dm_gcm3_1), main = "log(SLA)", xlab = "")
```

```{r}
par(mfrow = c(1, 2))
par(mar = c(2, 4, 2, 0.5))
hist(my_traits_effect$n_mgg_1 , main = "SLA", xlab = "")
hist(log(my_traits_effect$n_mgg_1), main = "log(SLA)", xlab = "")
```
'

```{r}
par(mfrow = c(1, 2))
par(mar = c(2, 4, 2, 0.5))
hist(my_traits_effect$p_mgg_1 , main = "SLA", xlab = "")
hist(log(my_traits_effect$p_mgg_1), main = "log(SLA)", xlab = "")
```


```{r}
resCWM <- functcomp(my_traits_effect, log(t(my_community) + 1), CWM.type = "all")
```
```{r}
par(mfrow = c(2, 2))
par(mar = c(3, 4, 2, 1))
plot(data_env$elev, resCWM$cfms_mgg_1, xlab = "moisture index",
ylab = "CWM SLA [log mm2/mg]", pch = 20)
plot(data_env$elev, resCWM$dm_gcm3_1, xlab = "moisture index",
ylab = "% of Shrubs", pch = 20)
plot(data_env$elev, resCWM$afe_mm2mg_1, xlab = "moisture index",
ylab = "% of Legumes", pch = 20)
plot(data_env$elev, resCWM$p_mgg_1, xlab = "moisture index",
ylab = "% of Therophytes", pch = 20)
```


```{r}
summary(lm(resCWM$afe_mm2mg_1 ~ elev*ca, data = data_env ))
```
```{r}

row.names(my_env) == row.names(resCWM)
my_env <- 
    data_env %>% 
    column_to_rownames("plot") %>% 
    select(-forest_type)
library(vegan)
rdaNEspain.all <- rda(resCWM ~ elev, data = data_env)
 # it is already loaded when you use the package 'FD'
rdaNEspain.all <- rda(resCWM ~ moisture.index * grazing, data = envxp)
plot(rdaNEspain.all, type = "n", scaling = "sites")
text(rdaNEspain.all, dis = "cn", scaling = "sites")
text(rdaNEspain.all, dis = "sp", scaling = "sites", col = "red")
```


```{r}
rdaNEspain0 <- rda(resCWM ~ 1, data = my_env)
rdaNEspain.all <- rda(resCWM ~ ca * elev, data = my_env)
ordistep(rdaNEspain0, scope = formula (rdaNEspain.all), direction = 'forward')
```

```{r}
RsquareAdj (rdaNEspain.all)$adj.r.square
```

